---
title: "Pb Hlth 244 proj 03"
author: "Adrienne Kwok"
date: "4/4/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(data.table)
library(ff)
library(dplyr)
library(ggplot2)
library(biglm)
library(broom)
```

```{r}
# grouping timeseries_us by date
# table of sum of number of cases per date
cases_bydate = timeseries_us %>% group_by(date) %>%
  summarise(sum_cases = sum(cases))

deaths_bydate = timeseries_us %>% group_by(date) %>%
  summarise(sum_deaths = sum(deaths))
  ## i have no idea why this group_by isn't working
```

```{r}
ggplot(subset(cases_bydate, !is.na(sum_cases)), aes(x = date, y = sum_cases)) +
  geom_line() + 
  labs(title = "Cases over time",
       subtitle = "in the US",
       y = "Cases",
       x = "Date")
```

```{r}
ggplot(subset(deaths_bydate, !is.na(sum_deaths)), aes(x = date, y = sum_deaths)) +
  geom_line() + 
  labs(title = "Deaths over time",
       subtitle = "in the US",
       y = "Deaths",
       x = "Date")
```

```{r}
# linear regression
# regressing number of cases on date and state
## tbh i'm not sure what's the diff btw "cases" and "active"... if only we had a dictionary
lm_cases = biglm(formula = cases ~ date + state,
       data = timeseries_us)
summary(lm_cases)
```


```{r}
# there are a lot of NAs in "active" column
lm_activecases = biglm(formula = active ~ date + state,
       data = timeseries_us)
summary(lm_activecases)
```

## Task 1: <p>

Forecast the number of daily cases and deaths: You may choose a cutoff date,
and use the numbers before that date as the training samples, and the numbers
after that date as the testing samples. Alternatively, you may choose a subset of
countries as the training samples, and the rest of countries as the testing samples.
```{r}
#Columns to work with: County, Province_State, Population, Weight, Date, Target, Target Value
dat <- fread(file = "data-proj-3a.csv", header=T)

dat_us = dat %>% filter(Country_Region == "US", Date > "2020-03-15") %>% 
  filter(TargetValue > 0)

dat_group = dat_us %>% group_by(Province_State, Date, Target) %>% summarize(total_TV = sum(TargetValue),
                                                                            total_pop = sum(Population) / 2)

  # data grouped by state, including US total
dat_state = dat_group %>% filter(Province_State != "") # data grouped by state only
dat_state$total_TV2 = sqrt(dat_state$total_TV) # adding transformed variable

dat_cases = dat_state %>% filter(Target == "ConfirmedCases") 
dat_cases = dat_cases %>% 
  filter(total_TV < (quantile(dat_cases$total_TV, .75) - quantile(dat_cases$total_TV, .25))*1.5)

dat_fat = dat_state %>% filter(Target == "Fatalities")
dat_fat = dat_fat %>% 
  filter(total_TV < (quantile(dat_fat$total_TV, .75) - quantile(dat_fat$total_TV, .25))*1.5)

#split into test and training by ... date ? First date: 2020/01/23 Last date: 2020/06/10
#midpoint --> 2020/04/01
# dat <- dat %>% filter( TargetValue > 100 & TargetValue < 1001) %>% mutate(log_tv = log(TargetValue))

set.seed(244)
dt_cases <-  sort(sample(nrow(dat_cases), nrow(dat_cases)*.75))
dt_fat <- sort(sample(nrow(dat_fat), nrow(dat_fat)*.75))

c_train<- dat_cases[dt_cases,] %>% data.frame()
c_test<- dat_cases[-dt_cases,] %>% data.frame()

f_train<- dat_fat[dt_fat,] %>% data.frame()
f_test<- dat_fat[-dt_fat,] %>% data.frame()

# c_train <- dat_cases %>% filter(Date < "2020-04-01")
# c_test <- dat_cases %>% filter(Date >= "2020-04-01")
# 
# f_train <- dat_fat %>% filter(Date < "2020-04-01")
# f_test <- dat_fat %>% filter(Date >= "2020-04-01")
```

Before going into building models... we should check for linear regression assumptions! 
- Relationship between X and Y is linear 
- Observations are independent of one another
- Normality: For any fixed value of X, Y is normally distributed
- Homoscedasticity: the variance of the residual is the same for any value X 
```{r}
dat %>% filter(Target == "Fatalities") %>% ggplot(aes(x = TargetValue)) + geom_histogram()

# for all countries
dat %>% filter(Target == "Fatalities") %>% ggplot(aes(x = Date, y = TargetValue)) + 
  geom_point()


# for US
dat_us %>% filter(Target == "Fatalities") %>% ggplot(aes(x = Date, y = TargetValue)) + 
  geom_point()

dat_cases %>% ggplot(aes(x = Date, y = TargetValue2)) + 
  geom_point()
```

```{r}
#no population
# for cases
lm_state_cases = lm(total_TV2 ~ Date + Province_State ,
      data = c_train)
glance(lm_state_cases)
tidy(lm_state_cases)

# for fatalities
lm_state_fat = lm(total_TV2 ~ Date + Province_State ,
      data = f_train)
glance(lm_state_fat)
tidy(lm_state_fat)
```

```{r}
c_test$pred_case <- predict(lm_state_cases, newdata = c_test)
c_test$resid <- c_test$total_TV2 - c_test$pred_case

f_test$pred_case <- predict(lm_state_fat, newdata = f_test)
f_test$resid <- f_test$total_TV2 - f_test$pred_case

#create scatterplot of original data values
c_test %>% ggplot(aes(x = pred_case, y = total_TV2)) + geom_point() 
c_test %>% ggplot(aes(x = pred_case, y = resid)) + geom_point() 

f_test %>% ggplot(aes(x = pred_case, y = total_TV2)) + geom_point() 
f_test %>% ggplot(aes(x = pred_case, y = resid)) + geom_point() 


library(MLmetrics)
RMSE((c_test$pred_case), (c_test$total_TV2))
#add predicted lines based on quadratic regression model
```





```{r}
library(ggplot2)
c_train %>%  ggplot(aes(x = TargetValue)) + geom_histogram()
c_test %>% ggplot(aes(x =  TargetValue)) + geom_histogram()
```

```{r}
library(broom)
#Let's build 2 linear regression models
c_lm <- biglm(TargetValue ~ Date + Weight  + factor(Country_Region),
              data = c_train, weight = Weight )
f_lm <- biglm(TargetValue ~ Date + Weight  + factor(Country_Region),
              data = f_train)

tidy(c_lm)  #should we keep Country_Region ?? 
tidy(f_lm)
#Use predict to test on the test sets then calculate residuals 
c_predict <- predict(c_lm, newdata = c_test)
f_predict <- predict(f_lm, newdata = f_test)

#Calculate residuals between observed and expected 
c_resid <- c_test$TargetValue - c_predict
f_resid <- f_test$TargetValue - f_predict

```


Let's plot some residuals 
```{r}
library(ggplot2)
c_df <- data.frame(daily_cases = c_test$TargetValue,
                         pred_cases = c_predict,
                         case_resid = c_resid)

f_df <- data.frame(daily_fat = f_test$TargetValue,
                         pred_fat = f_predict,
                         fat_resid = f_resid)

c_df %>% ggplot(aes(x = pred_cases, y = c_resid)) + 
  geom_point() + geom_hline(yintercept = 0, linetype = "dotted") + theme_bw()

f_df %>% ggplot(aes(x = pred_fat, y = f_resid)) + 
  geom_point() + geom_hline(yintercept = 0, linetype = "dotted") + theme_bw()
  
#They look weird L O L 

f_df %>% ggplot(aes(x = pred_fat, y = daily_fat)) + geom_poin
```







## Task 2 - 

What are the risk factors associated with the severity of illness from Covid-19 infection? For
example, profile of patients and lifestyle, social, co-morbidities, genetics, viral strain, etc.
influencers – how do these affect the probability of requiring hospitalization, ICU, ventilator,
mortality?

Looking only @ USA Data 
```{r}
mobility = fread("RDSC-07-30-Update/apple_mobility_trends/mobility-trends.csv", header=T) %>% 
  filter(region %in% state.name)

dat_us = dat %>% filter(Country_Region == "US")

world_case_death = fread("RDSC-07-30-Update/worldometer/worldometer-confirmed-cases-and-deaths-by-country-territory-or-conveyance.csv", header=T) %>% filter(country == "USA") # is a single row 

timeseries = fread("RDSC-07-30-Update/coronadatascraper/coronadatascraper-timeseries.csv", header=T) %>% 
  filter(country == "United States")
pop_days = fread("RDSC-07-30-Update/ECDC/current-data-on-the-geographic-distribution-of-covid-19-cases-worldwide.csv", header=T) %>% filter(countryterritorycode == "USA")

mobility3 = fread("RDSC-07-30-Update/google_mobility/us-mobility.csv", header=T)
mobility2_us = fread("RDSC-07-30-Update/descartes_labs/us-mobility-data.csv", header=T)

  # population counts, cases, deaths in US as a whole
cases_us = fread("RDSC-07-30-Update/USAFacts/confirmed-covid-19-cases-in-us-by-state-and-county.csv", header=T)
  # cases by state
deaths_us = fread("RDSC-07-30-Update/USAFacts/confirmed-covid-19-deaths-in-us-by-state-and-county.csv", header=T)
  # deaths by state
reduced_care = fread(
  "Indicators_of_Reduced_Access_to_Care_Due_to_the_Coronavirus_Pandemic_During_Last_4_Weeks.csv", header=T)
```






